refnx -- Neutron and X-ray reflectometry analysis in Python

### 2018-07-25 paper/manuscript.tex
PAPER: abstract

\begin{abstract}
The \refnx\ package is designed for object-oriented model-based analysis of neutron and X-ray reflectometry data in Python. It is cross platform, and has been tested on Linux, macOS, and Windows.
Model construction is modular, being composed from a series of components that each describe a subset of the interface, parameterised in terms of physically relevant parameters (volume fraction of a polymer, lipid area per molecule, etc). The model and data are used to create an objective, which is used to calculate residuals, log-likelihood, and log-prior probabilities of the system. Objectives are combined to perform co-refinement of multiple datasets, and mixed-area models. Prior knowledge of parameter values is encoded as probability distribution functions or bounds on all parameters in the system. Extra prior probability terms can be defined for Components, over and above those available from the Parameters alone. Algebraic Parameter constraints are available.
A choice of fitting approaches is available include least-squares (global and gradient-based optimizers) and a Bayesian approach using Markov Chain Monte Carlo to investigate the posterior distribution of the model parameters. The Bayesian approach is useful in examining parameter covariances, model selection, and variability in the resulting scattering length density profiles.
The package is designed to facilitate reproducible research; its use in \Jupyter\ notebooks (which includes a graphical user interface), and subsequent distribution of those notebooks as supporting information, permits straightforward reproduction of analyses.
\end{abstract}

@@

The \Structure\ object represents the interfacial model, comprising of individual \Component\ objects in series. Each \Component\ represents a subset of the interface and contains several physically relevant Parameters. The simplest \Component\ is a \Slab\, which has a uniform scattering length density (SLD), thickness, roughness, and volume fraction of solvent. The simplest models are simply a series of \Slab objects. More sophisticated components include \LipidLeaflet\ and \Spline. \Spline\ is designed for freeform modelling of an SLD profile using spline interpolation. In addition to its constituent parameters, each \Component\ can also contribute to the log-prior probability. This is useful when a \Component\ has a derived value, such as surface coverage, which is already known, and can be included as prior knowledge of the system. Each \Component\ has a \emph{slabs} property that represents one or more (micro)slices for its particular subset of the interface. A \Slab\ object has a single slice because it is a single thickness of uniform SLD. A \LipidLeaflet\ can be made of two slices (head/tail regions), but the \Spline\ has many thin slices approximating the smooth curve. Each of these slices has uniform SLD, the Nevot-Croce approach used to describe the roughness between them \cite{Nevot1980}; the user can create different roughness models by creating a `microsliced' \Component.

The \Structure\ object is used to construct a \ReflectModel\ object. This object is responsible for calculating the resolution smeared reflectivity of the \Structure, scaling the data, and adding a linear background (via the scale and background \Parameter objects). There are different types of smearing available: constant $\mathrm{d}Q/Q$, point-by-point resolution smearing read from the dataset of interest, or via a smearing probability kernel of arbitrary shape \cite{Nelson2014}. The constant $\mathrm{d}Q/Q$ and point-by-point smearing use Gaussian convolution, with $\mathrm{d}Q$ representing the full width half maximum (FWHM) of a Gaussian approximation to the instrument resolution function \cite{Well2005}.

>>>>

\begin{abstract}
\refnx\ is a model-based neutron and X-ray reflectometry data analysis package written in Python. It is cross platform, and has been tested on Linux, macOS, and Windows. Its graphical user interface is browser-based, through a \Jupyter\ notebook.
Model construction is modular, being composed from a series of components that each describe a subset of the interface, parameterised in terms of physically relevant parameters (volume fraction of a polymer, lipid area per molecule, etc). The model and data are used to create an objective, which is used to calculate residuals, log-likelihood, and log-prior probabilities of the system. Objectives are combined to perform co-refinement of multiple datasets, and mixed-area models. Prior knowledge of parameter values are encoded as probability distribution functions or bounds on all parameters in the system. Extra prior probability terms can be defined for Components, over and above those available from the Parameters alone. Algebraic Parameter constraints are available.
A choice of fitting approaches is available, including least-squares (global and gradient-based optimizers) and a Bayesian approach using Markov Chain Monte Carlo to investigate the posterior distribution of the model parameters. The Bayesian approach is useful in examining parameter covariances, model selection, and variability in the resulting scattering length density profiles.
The package is designed to facilitate reproducible research; its use in \Jupyter\ notebooks, and subsequent distribution of those notebooks as supporting information, permits straightforward reproduction of analyses.
\end{abstract}

@@

The \Structure\ object represents the interfacial model, comprising of individual \Component\ objects in series. Each \Component\ represents a subset of the interface and contains several physically relevant Parameters. The simplest \Component\ is a \Slab\, which has a uniform scattering length density (SLD), thickness, roughness, and volume fraction of solvent. The simplest models are simply a series of \Slab objects. More sophisticated components include \LipidLeaflet\ and \Spline. \Spline\ is designed for freeform modelling of an SLD profile using spline interpolation. In addition to its constituent parameters, each \Component\ can also contribute to the log-prior probability. This is useful when a \Component\ has a derived value, such as surface excess, which is already known, and can be included as prior knowledge of the system. Each \Component\ has a \emph{slabs} property that represents one or more (micro)slices for its particular subset of the interface. A \Slab\ object has a single slice because it is a single thickness of uniform SLD. A \LipidLeaflet\ can be made of two slices (head/tail regions), but the \Spline\ has many thin slices approximating the smooth curve. Each of these slices has uniform SLD, the Nevot-Croce approach used to describe the roughness between them \cite{Nevot1980}; the user can create different roughness models by creating a `microsliced' \Component.

The \Structure\ object is used to construct a \ReflectModel\ object. This object is responsible for calculating the resolution smeared reflectivity of the \Structure, scaling the data, and adding a $Q$-independent linear background (via the scale and background \Parameter\ objects). There are different types of smearing available: constant $\mathrm{d}Q/Q$, point-by-point resolution smearing read from the dataset of interest, or via a smearing probability kernel of arbitrary shape \cite{Nelson2014}. The constant $\mathrm{d}Q/Q$ and point-by-point smearing use Gaussian convolution, with $\mathrm{d}Q$ representing the full width half maximum (FWHM) of a Gaussian approximation to the instrument resolution function \cite{Well2005}.
